### 2018-04-23 Presentations/Milestone1/presentation.tex
Add Template for first presentation

 \documentclass{beamer}

\usetheme{MagdeburgFIN}
\usefonttheme{structurebold}
\usepackage{graphicx}
\usepackage{float}
\usepackage{url}
\usepackage{pdfpages}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\title{Scientific Team Project}
\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising}
\date{\today}
\institute{Otto von Guericke University, Magdeburg}

\begin{document}

\begin{frame}[plain]
 \titlepage
\end{frame}

\section[Agenda]{}
\begin{frame}
\frametitle{Agenda}
\tableofcontents
\end{frame}

\begin{frame}
 \frametitle{Thank you for your attention!}
\end{frame}
\end{document}