### 2018-04-23 Presentations/Milestone1/presentation.tex
Added Intro

 \documentclass{beamer}

\usetheme{MagdeburgFIN}
\usefonttheme{structurebold}
\usepackage{graphicx}
\usepackage{float}
\usepackage{url}
\usepackage{pdfpages}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\title{Scientific Team Project}
\subtitle{Thread Pool in GeckoDB/BOLSTER}
\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
\date{\today}
\institute{Otto von Guericke University, Magdeburg}

%Topic, schedule, and team presentation & first results of literature research

\begin{document}

\begin{frame}[plain]
 \titlepage
\end{frame}

\section[Agenda]{}
	\begin{frame}
	\frametitle{Agenda}
	\tableofcontents
	\end{frame}

\section{Introduction}
	\begin{frame}
		\frametitle{Introduction}
		\begin{itemize}
			\item \emph{In computer programming, a thread pool is a software design pattern for achieving concurrency of execution in a computer program.}
			\item Pool of Threads
			\item Reduces overhead of creating new threads by reusing existing threads.
		\end{itemize}
	\end{frame}
	\begin{frame}
		\frametitle{Current Situation}
		\begin{itemize}
			\item Operations are spawning dozens of threads, which are not handled and managed..
			\item After some operations profilers dies.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Desirable Situation}
		\begin{itemize}
			\item Reuse of Threads with thread pools.
			\item Limited amount of threads, which are handled correctly.
		\end{itemize}
	\end{frame}

\begin{frame}
 \frametitle{Thank you for your attention!}
\end{frame}
\end{document}