### 2018-04-23 Presentations/Milestone1/presentation.tex
Reworked parts of presentation1

 \documentclass{beamer}

\usetheme{MagdeburgFIN}
\usefonttheme{structurebold}
\usepackage{graphicx}
\usepackage{float}
\usepackage{url}
\usepackage{pdfpages}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\title{Scientific Team Project}
\subtitle{Thread Pool in GeckoDB/BOLSTER}
\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
\date{\today}
\institute{Otto von Guericke University, Magdeburg}

%Topic, schedule, and team presentation & first results of literature research

\begin{document}

\begin{frame}[plain]
 \titlepage
\end{frame}

\section[Agenda]{}
	\begin{frame}
	\frametitle{Agenda}
	\tableofcontents
	\end{frame}

\section{Motivation}
	\begin{frame}
		\frametitle{Motivation}
		\begin{center}
			\includegraphics[width=0.6\textwidth]{img/bigpicture.png}
		\end{center}
	\end{frame}

\section{Introduction}
	\begin{frame}
		\frametitle{Introduction}
		\begin{itemize}
			\item \emph{In computer programming, a thread pool is a software design pattern for achieving concurrency of execution in a computer program.}
			\item Pool of Threads to distribute parallel tasks
			\item Reduces overhead of creating new threads by reusing existing threads.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Current Situation}
		\begin{itemize}
			\item Operations are spawning dozens of threads, which are not handled and managed..
			\item Heavy overhead of creating and removing threads
			\item After some operations profiler dies (too many threads).
			\item Debugging/Profiling not possible due to number of threads 
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Desirable Situation}
		\begin{itemize}
			\item Reuse of Threads with thread pools.
			\item Reduce overhead of thread management
			\item Improve maintainability of parallel code
			\item Defy competition to existing thread pool libs
		\end{itemize}
	\end{frame}

\section{Planning}

\begin{frame}
	\frametitle{Team Roles}
	\begin{figure}
		\includegraphics[width=0.7\linewidth]{img/TeamRoles}
	\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Schedule}
	\includegraphics[width=1.0\textwidth]{img/schedule.png}
\end{frame}

\section{Literature Research}
    \begin{frame}
        \frametitle{Literature Research}
        \begin{itemize}
            \item Xu, Dongping, and Brett Bode. \emph{"Performance study and dynamic optimization design for thread pool systems".} Diss. United States. Department of Energy. Office of Science, 2004.
            \item Ling, Yibei, Tracy Mullen, and Xiaola Lin. \emph{Analysis of optimal thread pool size.} ACM SIGOPS Operating Systems Review 34.2 (2000): 42-55.
            \item Kang, DongHyun, et al. \emph{Prediction-based dynamic thread pool scheme for efficient resource usage.} Computer and Information Technology Workshops, 2008. CIT Workshops 2008. IEEE 8th International Conference on. IEEE, 2008.
        \end{itemize}
    \end{frame}

    \begin{frame}
        \frametitle{Thank you for your attention!}
    \end{frame}
\end{document}