### 2018-05-09 Presentations/Milestone2/presentation.tex
Begin milestone 2 presentation

 \documentclass{beamer}

\usetheme{MagdeburgFIN}
\usefonttheme{structurebold}
\usepackage{graphicx}
\usepackage{float}
\usepackage{url}
\usepackage{pdfpages}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\title{Thread Pool in GeckoDB/BOLSTER}
\subtitle{Milestone II: Concept \& additional literature research}
\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
\date{\today}
\institute{Otto von Guericke University, Magdeburg}

% Milestone II: Concept & additional literature research

\begin{document}

\begin{frame}[plain]
 \titlepage
\end{frame}

\section[Agenda]{}
	\begin{frame}
	\frametitle{Agenda}
	\tableofcontents
	\end{frame}

\section{Previously...}
\begin{frame}
	\frametitle{Previously in Thread Pool in GeckoDB/BOLSTER}
	\begin{itemize}
		\item Operations are spawning dozens of threads, which are not handled and managed
		\item Heavy overhead of creating and removing threads
		\item After some operations profiler dies (too many threads)
	\end{itemize}
	\begin{center}
		\includegraphics[width=0.3\textwidth]{img/thumbs_down.png}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Previously in Thread Pool in GeckoDB/BOLSTER}
	\begin{itemize}
		\item Reuse of Threads with thread pools.
		\item Reduce overhead of thread management
		\item Improve maintainability of parallel code
	\end{itemize}
	\begin{center}
		\includegraphics[width=0.3\textwidth]{img/thumbs_up.png}
	\end{center}
\end{frame}

\section{Concept and Design}
\begin{frame}
	\frametitle{Concept}
	\begin{itemize}
		\item Thread pool with runtime fixed amount of threads and lock free priority queue
		\item Tasks are not preemtable and independent
		\item Array of tasks passed to an enqueue function, which waits until all tasks are finished
		\item Threads without tasks idle until they get a new one
		\item Monitoring/managing state and assigned task of each thread
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Design}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{img/pool_queue.png}
	\end{center}
\end{frame}

\section{Current Progress}

\section{Next Steps}

\section{Additional Literature Research}
    \begin{frame}
        \frametitle{Literature Research}
        \begin{itemize}
            \item 
        \end{itemize}
    \end{frame}

    \begin{frame}
        \frametitle{Thank you for your attention!}
    \end{frame}
\end{document}