### 2018-06-07 Presentations/Milestone3/presentation.tex
Fred did some awesome stuff

 \documentclass{beamer}

\usetheme{MagdeburgFIN}
\usefonttheme{structurebold}
\usepackage{graphicx}
\usepackage{wrapfig,lipsum}
\usepackage{float}
\usepackage{url}
\usepackage{pdfpages}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\title{Thread Pool in GeckoDB/BOLSTER}
\subtitle{Milestone III: Implementation \& evaluation setup}
\author{Johann Wagner, Johannes Wünsche, Marten Wallewein-Eising, Robert Jenderise, Benedikt Zeilinger}
\date{\today}
\institute{Otto von Guericke University, Magdeburg}

\begin{document}

\begin{frame}[plain]
 \titlepage
\end{frame}

\section[Agenda]{}
	\begin{frame}
	\frametitle{Agenda}
	\tableofcontents
	\end{frame}

\section{Previously...}
\begin{frame}
	\begin{center}
		\huge Previously...
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Previously on Thread Pool in GeckoDB/BOLSTER}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{img/pool_queue.png}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Previously on Thread Pool in GeckoDB/BOLSTER}
	\begin{columns}
		\begin{column}{0.48\textwidth}
			\begin{itemize}
				\item Problem: waiting for a task to complete could lead to a
				race condition
				\item This was caused by the uncertainty of a thread\_task
				during transition between the priority queue and the
				thread pool
				\item Solution: slotmap with an atomic counter
			\end{itemize}
		\end{column}
		\begin{column}{0.48\textwidth}
			\includegraphics[width=1.0\textwidth]{waitingconcept.png}
		\end{column}
	\end{columns}
\end{frame}

\section{Implementation}
\begin{frame}
	\begin{center}
		\huge Implementation
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Design of Thread Pool System}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{img/pool_structure.png}
	\end{center}
	\begin{itemize}
		\item Thread pool system contains priority queue and thread pool itself
		\item Performance monitoring as optional, external component
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Implementation}
	We...
	\begin{itemize}
		\item Implemented task enqueueing and execution
		\item Implemented performance monitoring as external component
		\item Added gtest framework to project and wrote tests for pool and queue
		\item Started writing performance tests and measures for comparison against baseline ???
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Open issues}
	\begin{itemize}
		\item gtest still has some problems due to itself being implemented in C++
		\item some race condition may still occur
	\end{itemize}
\end{frame}

\section{Evaluation Setup}
\begin{frame}
	\begin{center}
		\huge Evaluation Setup
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Task Statistics}
	\begin{center}
		\begin{tabular}{ c c }
			\hline
			\textbf{Statistic name}&\textbf{Description}\\
			\hline
			enqueue\_time & Time when the task was enqueued \\
			execution\_time & Time when the task execution begins \\
			complete\_time & Time when the task execution is finished \\
			\hline
		\end{tabular}
		\label{tab2}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Thread Statistics}
	\begin{center}
		\begin{tabular}{ c c }
			\hline
			\textbf{Statistic name}&\textbf{Description}\\
			\hline
			waiting\_time & Time the thread spend waiting (ms) \\
			busy\_time & Time the thread spend executing tasks (ms)\\
			task\_count & Number of tasks the thread has executed \\
			\hline
		\end{tabular}
		\label{tab3}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Thread Pool Statistics}
	\begin{center}
		\begin{tabular}{ c c }
			\hline
			\textbf{Statistic name}&\textbf{Description}\\
			\hline
			working\_time & Time the thread pool runs (ms) \\
			task\_complete\_count & Number of completed tasks \\
			task\_enqueued\_count & Number of enqueued tasks \\
			avg\_complete\_time & Average time span from enqueueing until \\
			& execution of tasks \\
			avg\_wait\_time & Average time of tasks spend in the queue \\
			\hline
		\end{tabular}
		\label{tab4}
	\end{center}
\end{frame}

\section{Paper Progress}
\begin{frame}
	\begin{center}
	\huge Paper Progress \normalsize	
	\end{center}
\end{frame}

\begin{frame}
	\begin{itemize}
		\item Introduction and Preliminaries mostly completed
		\item Implementation described, including our used concepts and basic ideas how to build and operate the pool
		\item Still required is the Evaluation(since we still need to perform it), the description of our testing environment and our final conclusion
	\end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Thank you for your attention!}
 	\includegraphics[width=\textwidth]{img/important.jpg}
\end{frame}

\end{document}